---
title: "PSI_Assignment 1: Digital Portfolio: Prepare and Explore Parts"
author: Na Li (D19125334)
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)

needed_packages <- c("VIM", "tidyverse","pastecs", "ggplot2", "semTools", "psych", "FSA", "car", "effectsize", "coin", "rstatix", "sjstats", "userfriendlyscience", "stats", "foreign", "gmodels")   
# Extract not installed packages
not_installed <- needed_packages[!(needed_packages %in% installed.packages()[, "Package"])]    
# Install not installed packages
if(length(not_installed)) install.packages(not_installed) 

library(pastecs) #For creating descriptive statistic summaries
library(ggplot2) #For creating histograms with more detail than plot
library(semTools) #For skewness and kurtosis
library(psych)  #For descriptive functions
library(FSA) #For percentage
library(car) # For Levene's test for homogeneity of variance 
library(effectsize) #To calculate effect size for t-test
library(VIM)
library(tidyverse)
library(coin) # For Wilcox test (non-parametric)
library(rstatix) # For calculating effect size
library(sjstats) #calculate effect size for t-test
library(userfriendlyscience)
library(stats)
library(foreign) # open SPSS file, I may not use that.
library(gmodels) #For creating histograms with more detail than plot
```

# Understanding Dataset

This dataset is Secondary School student Performance in secondary education of two Portuguese schools, in this secion, 
1. Merged data set from two csv files (student-mat.csv (Math course) and student-por.csv (Portuguese language course) datasets), account the number of data set as a sample data. 
2. Visualising Quantitative Data
3. Visualising Qualitative Data


```{r initial dataset, echo=TRUE}

getwd()

stdMathData=read.table("student-mat.csv",sep=";",header=TRUE) #read student 
stdPorlagData=read.table("student-por.csv",sep=";",header=TRUE)

stdMergeData=merge(stdMathData,stdPorlagData,by=c("school","sex","age","address","famsize","Pstatus","Medu","Fedu","Mjob","Fjob","reason","nursery","internet"))

#total number of the dataset
print(nrow(stdMergeData)) # 382 records/rows/cases
#number of students math course -> 395 records
nrow(stdMathData)
#number of students for Portuguese language course -> 649 records
nrow(stdPorlagData)

colnames(stdMathData) <- tolower(colnames(stdMathData))
colnames(stdPorlagData) <- tolower(colnames(stdPorlagData))
colnames(stdMergeData) <- tolower(colnames(stdMergeData))

write.csv(x=stdMergeData, file="student_merged.csv")

#show variables in three datasets
colnames(stdMathData)
colnames(stdPorlagData)
colnames(stdMergeData)

```
## Qualitative Data
```{r, echo=TRUE}
# Looking at interest variables in two Samples: sex, mother jobs, father job, reason
# Qualitative data 

stdMathData$sex <- as.factor(stdMathData$sex)
stdPorlagData$sex <- as.factor(stdPorlagData$sex)
stdMergeData$sex <- as.factor(stdMergeData$sex)

summary(stdMathData$sex)
summary(stdPorlagData$sex)
summary(stdMergeData$sex)

```

```{r, echo=TRUE}
# Looking at interest variables in two Samples: sex, mother jobs, father job, reason
# Qualitative data 

stdMathData$mjob <- as.factor(stdMathData$mjob)
stdPorlagData$mjob <- as.factor(stdPorlagData$mjob)

summary(stdMathData$mjob)
summary(stdPorlagData$mjob)
```
```{r descripting fjob, echo=TRUE}
# Looking at interest variables in two Samples: sex, mother jobs, father job, reason
# Qualitative dataset 
stdMathData$fjob <- as.factor(stdMathData$fjob)
stdPorlagData$fjob <- as.factor(stdPorlagData$fjob)

summary(stdMathData$fjob)
summary(stdPorlagData$fjob)
```
```{r descripting address, echo=TRUE}
# Looking at interest variables in two Samples: address
# Qualitative data 
stdMathData$address <- as.factor(stdMathData$address)
stdPorlagData$address <- as.factor(stdPorlagData$address)
# Student data for Math course: reason to choose this school 
summary(stdMathData$address)
```
``` {r descripting address-summary, echo=TRUE}
# Student data for Portuguese language course: reason to choose this school 
summary(stdPorlagData$address)
```
``` {r descripting famsup, echo=TRUE}
# Looking at interest variables in two Samples: family educational support 
# Qualitative data 
stdMathData$famsup <- as.factor(stdMathData$famsup)
stdPorlagData$famsup <- as.factor(stdPorlagData$famsup)
# Math course
str(stdMathData$famsup)
summary(stdMathData$famsup)
```
``` {r descripting summary famsup, echo=TRUE}
# Looking at interest variables in two Samples: family educational support 
# Qualitative data 
# Portuguese language course
str(stdPorlagData$famsup)
summary(stdPorlagData$famsup)
```
``` {r descripting romantic, echo=TRUE}
# Looking at interest variables in two Samples: with a romantic relationship 
# Qualitative data 
stdMathData$romantic <- as.factor(stdMathData$romantic)
stdPorlagData$romantic <- as.factor(stdPorlagData$romantic)
# Math course
str(stdMathData$romantic)
summary(stdMathData$romantic)
```

``` {r descripting romantic for Portuguese, echo=TRUE}
# Looking at interest variables in two Samples: with a romantic relationship 
# Qualitative data 
# Portuguese language course
str(stdPorlagData$romantic)
summary(stdPorlagData$romantic)
```
## Quantitative Data
Looking at interest quantitativ data of variables in two samples: traveltime, studytime, G1 - G3 (grade 1 to grade 3),
### Measure of Central Tendency:mean, median, mode, 
```{r descripting sample: mean, echo=TRUE}
#traveltime: math course
mean(stdMathData$traveltime)
#traeltime: portuguese language course
mean(stdPorlagData$traveltime) 
```
```{r descripting sample: median, echo=TRUE}
#traveltime: math course
median(stdMathData$traveltime)
#traeltime: portuguese language course
median(stdPorlagData$traveltime) 
```

```{r get mode function, echo=TRUE}
#traveltime: math course
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}

getmode(stdMathData$traveltime)
#traeltime: portuguese language course
getmode(stdPorlagData$traveltime) 
``` 

```{r descripting sample: traveltime, echo=TRUE}
#traveltime: math course
summary(stdMathData$traveltime)
#traeltime: portuguese language course
summary(stdPorlagData$traveltime)
``` 
```{r descripting sample: studytime, echo=TRUE}
#studytime: math course
summary(stdMathData$studytime)
#studytime: portuguese language course
summary(stdPorlagData$studytime)
``` 
```{r descripting studytime: mode, echo=TRUE}
#studytime: math course
getmode(stdMathData$studytime)
#studytime: portuguese language course
getmode(stdPorlagData$studytime) 
```

```{r descripting g1, echo=TRUE}
#studytime: math course
summary(stdMathData$g1)
#studytime: portuguese language course
summary(stdPorlagData$g1)
``` 

```{r descripting g1: mode, echo=TRUE}
#G1: math course
getmode(stdMathData$g1)
#G1: portuguese language course
getmode(stdPorlagData$g1) 
```

```{r descripting g2, echo=TRUE}
#studytime: math course
summary(stdMathData$g2)
#studytime: portuguese language course
summary(stdPorlagData$g2)
``` 

```{r descripting g2: mode, echo=TRUE}
#G2: math course
getmode(stdMathData$g2)
#G2: portuguese language course
getmode(stdPorlagData$g2) 
```

```{r descripting g3, echo=TRUE}
#studytime: math course
summary(stdMathData$g3)
#studytime: portuguese language course
summary(stdPorlagData$g3)
``` 

```{r descripting g3: mode, echo=TRUE}
#G3: math course
getmode(stdMathData$g3)
#G3: portuguese language course
getmode(stdPorlagData$g3) 
```
### Measure of Dispersion (descriptive statistics): range, quantiles, standard deviation (sd), IQR
```{r descripting g1g2g3: SD, echo=TRUE}
#G1: math course
round(sd(as.double(stdMathData$g1)),2)
#G1: portuguese language course
round(sd(as.double(stdPorlagData$g1)),2)
```
```{r descripting g2: SD, echo=TRUE}
#G2: math course
round(sd(as.double(stdMathData$g2)),2)
#G2: portuguese language course
round(sd(as.double(stdPorlagData$g2)),2)
```
```{r descripting g3: SD, echo=TRUE}
#G3: math course
sdMathData_g3 <- sd(as.double(stdMathData$g3))
round(sdMathData_g3, 2)
#G3: portuguese language course
sdPorlagData_g3 <- sd(as.double(stdPorlagData$g3))
round(sdPorlagData_g3,2)

```

```{r descripting g1: range, quantile and Interquartile Range, echo=TRUE}
#math course
range(as.double(stdMathData$g1))
quantile(as.double(stdMathData$g1))
IQR(as.double(stdMathData$g1))
var(as.double(stdMathData$g1))
 
#portuguese language course
range(as.double(stdPorlagData$g1))
quantile(as.double(stdPorlagData$g1))
IQR(as.double(stdPorlagData$g1))
var(as.double(stdPorlagData$g1))
```
```{r descripting g2: range, quantile and Interquartile Range, echo=TRUE}
#math course
range(as.double(stdMathData$g2))
quantile(as.double(stdMathData$g2))
IQR(as.double(stdMathData$g2))
var(as.double(stdMathData$g2))
 
#portuguese language course
range(as.double(stdPorlagData$g2))
quantile(as.double(stdPorlagData$g2))
IQR(as.double(stdPorlagData$g2))
var(as.double(stdPorlagData$g2))
```
```{r descripting g3: range, quantile and Interquartile Range, echo=TRUE}
#math course
range(as.double(stdMathData$g3))
quantile(as.double(stdMathData$g3))
IQR(as.double(stdMathData$g3))
var(as.double(stdMathData$g3))
 
#Portuguese language course
range(as.double(stdPorlagData$g3))
quantile(as.double(stdPorlagData$g3))
IQR(as.double(stdPorlagData$g3))
var(as.double(stdPorlagData$g3))
```
### Visuallsing Quantitative Data
```{r plot G3, echo=TRUE}

gg_g3 <- ggplot(stdMathData, aes(x=stdMathData$g3))
gg_g3 <- gg_g3 + labs(x="G3")
gg_g3 <- gg_g3 + geom_histogram(bins = 30,aes(y=..density.., color=sex, fill=..count..)) 
   #geom_histogram(binwidth=2, colour="black", aes(y=..density.., fill=..count..))
       
  
#gg_g3 <- gg_g3 + scale_fill_gradient("Count", low="#DCDCDC", high="#7C7C7C")

gg_g3 <- gg_g3 + stat_function(fun=dnorm, color='red',args=list(mean=mean(stdMathData$g3, na.rm=TRUE), sd=sd(stdMathData$g3, na.rm=TRUE)))

gg_g3

```

```{r plot absences-Math, echo=TRUE}

ggplot(stdMathData, aes(x=stdMathData$absences)) + 
         labs(x="math: absences") +
         geom_histogram(bins = 30,aes(y=..density.., fill=..count..)) +
         stat_function(fun=dnorm, color='red',args=list(mean=mean(stdMathData$absences, na.rm=TRUE), sd=sd(stdMathData$absences, na.rm=TRUE)))

```
```{r plot absences-Portuguese, echo=TRUE}

gg_st <- ggplot(stdPorlagData, aes(x=stdPorlagData$absences))
gg_st <- gg_st + labs(x="Portuguese language: absences")
gg_st <- gg_st + geom_histogram(bins = 30,aes(y=..density.., fill=..count..)) 
gg_st <- gg_st + stat_function(fun=dnorm, color='red',args=list(mean=mean(stdPorlagData$absences, na.rm=TRUE), sd=sd(stdPorlagData$absences, na.rm=TRUE)))

gg_st

```

```{r plot studytime , echo=TRUE}
#pie chart
mytable <- table(stdMathData$studytime)
lbls <- paste(names(mytable), "\n", mytable, sep="")
pie(mytable, labels = lbls,
   main="Math: Weekly Study time \n 1 - <2 hours, 2 - 2 to 5 hours, 3 - 5 to 10 hours, or 4 - >10 hours")

# Pie Chart from data frame with Appended Sample Sizes
mytable_por <- table(stdPorlagData$studytime)
lbls <- paste(names(mytable_por), "\n", mytable_por, sep="")
pie(mytable, labels = lbls,
   main="Prtuguese language: Weekly Study time \n 1 - <2 hours, 2 - 2 to 5 hours, 3 - 5 to 10 hours, or 4 - >10 hours")
```

```{r plot g1g2g3 math , echo=TRUE}
gg_math <- ggplot(stdMathData, aes(x=stdMathData$g1))
gg_math <- gg_math + labs(x="Math:G1-Red, G2-Blue, G3-Green", y="Frequency")
gg_math <- gg_math + stat_function(fun=dnorm, color='red',args=list(mean=mean(stdMathData$g1, na.rm=TRUE), sd=sd(stdMathData$g1, na.rm=TRUE)))

gg_math <- gg_math + stat_function(fun=dnorm, color='blue',args=list(mean=mean(stdMathData$g2, na.rm=TRUE), sd=sd(stdMathData$g2, na.rm=TRUE)))


gg_math <- gg_math + stat_function(fun=dnorm, color='green',args=list(mean=mean(stdMathData$g3, na.rm=TRUE), sd=sd(stdMathData$g3, na.rm=TRUE)))

gg_math

```
```{r , echo=TRUE}
gg_porlag <- ggplot(stdPorlagData, aes(x=stdPorlagData$g1))
gg_porlag <- gg_porlag + labs(x="Portuguese language:G1-Red, G2-Blue, G3-Green", y="Frequency")
gg_porlag <- gg_porlag + stat_function(fun=dnorm, color='red',args=list(mean=mean(stdPorlagData$g1, na.rm=TRUE), sd=sd(stdPorlagData$g1, na.rm=TRUE)))

gg_porlag <- gg_porlag + stat_function(fun=dnorm, color='blue',args=list(mean=mean(stdPorlagData$g2, na.rm=TRUE), sd=sd(stdPorlagData$g2, na.rm=TRUE)))


gg_porlag <- gg_porlag + stat_function(fun=dnorm, color='green',args=list(mean=mean(stdPorlagData$g3, na.rm=TRUE), sd=sd(stdPorlagData$g3, na.rm=TRUE)))

gg_porlag

```

```{r plot G1 by sex group,echo=TRUE}

female <- filter(stdMathData, sex == "F")
male <- filter(stdMathData, sex=="M")

gg_g1 <- ggplot(stdMathData, aes(x=g1, fill=sex, colour=sex)) + geom_density(alpha = 0.1)
gg_g1 <- gg_g1 + labs(x="Math:Normality G1")

gg_g1 <- gg_g1 + stat_function(data = female, geom = "point", fun=dnorm, args=list(mean=mean(female$g1, na.rm=TRUE), sd=sd(female$g1, na.rm=TRUE)))

gg_g1 <- gg_g1 + stat_function(data = male, geom = "point", fun=dnorm, args=list(mean=mean(male$g1, na.rm=TRUE), sd=sd(male$g1, na.rm=TRUE)))

gg_g1

```
```{r plot G1 by sex group for Portuguese, echo=TRUE}
library(ggplot2)

female <- filter(stdPorlagData, stdPorlagData$sex=="F")
male <- filter(stdPorlagData, stdPorlagData$sex=="M")

gg_g1 <- ggplot(stdPorlagData, aes(x=g1, fill=sex, colour=sex)) + geom_density(alpha = 0.1)
gg_g1 <- gg_g1 + labs(x="Portuguese language: Normality G1")

gg_g1 <- gg_g1 + stat_function(data = female, geom = "point", fun=dnorm, args=list(mean=mean(female$g1, na.rm=TRUE), sd=sd(female$g1, na.rm=TRUE)))

gg_g1 <- gg_g1 + stat_function(data = male, geom = "point", n=200, fun=dnorm, args=list(mean=mean(male$g1, na.rm=TRUE), sd=sd(male$g1, na.rm=TRUE)))

gg_g1

```

```{r qq G3 , echo=TRUE}
#qq-plot
qqnorm(stdMathData$g3)
qqline(stdMathData$g3, col=2)
```

```{r statistic summary G3, echo=TRUE}
library(pastecs)
pastecs::stat.desc(stdMathData$g3, basic=F)
``` 
```{r tpskew&tpkurt g3, echo=TRUE}
tpskew<-semTools::skew(stdMathData$g3)
tpkurt<-semTools::kurtosis(stdMathData$g3)
tpskew[1]/tpskew[2]
tpkurt[1]/tpkurt[2]
```
```{r}
ztpcoiss<- abs(scale(stdMathData$g3))
FSA::perc(as.numeric(ztpcoiss), 1.96, "gt")
```
```{r}
FSA::perc(as.numeric(ztpcoiss), 3.29, "gt")
```

```{r plot G2 by sex group, echo=TRUE}
gg_g2 <- ggplot(stdMathData, aes(x=stdMathData$g2))
gg_g2 <- gg_g2 + labs(x="G2")
gg_g2 <- gg_g2 + geom_histogram(bins = 30,aes(y=..density.., fill=..count..,color=stdMathData$sex))
#gg_g2 <- gg_g2 + scale_fill_gradient("Count", low="#DCDCDC", high="#7C7C7C")

gg_g2 <- gg_g2 + stat_function(fun=dnorm, color="red",args=list(mean=mean(stdMathData$g2, na.rm=TRUE), sd=sd(stdMathData$g2, na.rm=TRUE)))

gg_g2

```

```{r plot G1 without group , echo=TRUE}
gg_g1 <- ggplot(stdMathData, aes(x=g1))
gg_g1 <- gg_g1 + labs(x="G1: Math")
gg_g1 <- gg_g1 + geom_histogram(binwidth=2, colour="black", aes(y=..density.., fill=..count..))
gg_g1 <- gg_g1 + scale_fill_gradient("Count", low="#DCDCDC", high="#7C7C7C")

gg_g1 <- gg_g1 + stat_function(fun=dnorm, color="red",args=list(mean=mean(stdMathData$g1, na.rm=TRUE), sd=sd(stdMathData$g1, na.rm=TRUE)))

gg_g1
```

``` {r , echo=TRUE}
library(ggplot2)
gg_g1 <- ggplot(stdMathData, aes(x=stdMathData$g1))
gg_g1 <- gg_g1 + labs(x="G1: Math")
gg_g1 <- gg_g1 + geom_histogram(bins = 30,aes(y=..density.., color=sex, fill=..count..))
gg_g1 <- gg_g1 + stat_function(fun=dnorm, color="red",args=list(mean=mean(stdMathData$g1, na.rm=TRUE), sd=sd(stdMathData$g1, na.rm=TRUE)))

gg_g1
```

```{r,echo=TRUE}
gg_g1 <- ggplot(stdPorlagData, aes(x=g1))
gg_g1 <- gg_g1 + labs(x="G1: Portuguese language")
gg_g1 <- gg_g1 + geom_histogram(binwidth=2, colour="black", aes(y=..density.., fill=..count..))
gg_g1 <- gg_g1 + scale_fill_gradient("Count", low="#DCDCDC", high="#7C7C7C")

gg_g1 <- gg_g1 + stat_function(fun=dnorm, color="red",args=list(mean=mean(stdPorlagData$g1, na.rm=TRUE), sd=sd(stdPorlagData$g1, na.rm=TRUE)))

gg_g1
```

``` {r ,echo=TRUE}
gg_g1 <- ggplot(stdPorlagData, aes(x=stdPorlagData$g1))
gg_g1 <- gg_g1 + labs(x="G1: Portuguese language")
gg_g1 <- gg_g1 + geom_histogram(bins = 30,aes(y=..density.., color=sex, fill=..count..)) 
gg_g1 <- gg_g1 + stat_function(fun=dnorm, color="red",args=list(mean=mean(stdPorlagData$g1, na.rm=TRUE), sd=sd(stdPorlagData$g1, na.rm=TRUE)))

gg_g1
```
### Missing Data
```{r exploring missing data, echo=TRUE}
variableint <- c("g1", "g2", "g3", "studytime", "traveltime", "health", "absences", "famrel", "freetime")
subdata_math <-stdMathData[variableint]
subdata_por <- stdPorlagData[variableint]
summary(subdata_math)
summary(subdata_por)
```

```{r show missing data: math course, echo=TRUE}
#Create and inspect patterns of missingness
res_math<-summary(VIM::aggr(subdata_math, sortVar=TRUE))$combinations

#Output the most common combinations
head(res_math[rev(order(res_math[,2])),])
```

```{r show missing data: Portuguese language course, echo=TRUE}
#Create and inspect patterns of missingness
res_portuggese<-summary(VIM::aggr(subdata_por, sortVar=TRUE))$combinations

#Output the most common combinations
head(res_portuggese[rev(order(res_portuggese[,2])),])
```
The results show that the variables of interests, no missing data.
 

# Explore Data
## descriptive statistics
```{r Generate Summary Statistics, echo=TRUE}
# Center Tendency
variableint <- c("g1", "g2", "studytime", "traveltime","absences")
subdata_math <-stdMathData[variableint]
subdata_por <- stdPorlagData[variableint]
summary(subdata_math)
summary(subdata_por)

## summary variability 
pastecs::stat.desc(stdMathData[variableint], basic=F)
pastecs::stat.desc(stdPorlagData[variableint], basic=F)
```


## Assessing a variable for Normality
Step 1: Assessing by plots
Histogram with normal curve 
QQ-plot checking if it can be normalised 
Step 2:  Summary Statistics including measure of skewness and kurtosis. 
Step 3: Review statistics by plotting to check whether it close to normal data. 
Step 4: Report correct statistics depend on the assessment and show the result whether the data can follow the normal distribution.

``` {r normality:histogram, echo=TRUE}
# histogram for math course sample

#study time
histogr_stdtime <- ggplot(stdMathData, aes(x=studytime)) + 
                   labs(x="Weekly Study Time for Math Course") + 
                   geom_histogram(bins=8, colour="blue", aes(y= ..density.., fill=..count..)) +
                   scale_fill_gradient("count", low="#132B43", high="#56B1F7") +
                   stat_function(fun=dnorm, color="red", args=list(mean=mean(stdMathData$studytime, na.rm=TRUE), sd=sd(stdMathData$studytime, na.rm=TRUE)))

histogr_stdtime
boxplot(stdMathData$studytime,
        main = "Math course: Weekly Study Time",
        col = "orange",
        border = "brown",
        horizontal = TRUE)

# absences
histogr_ab <- ggplot(stdMathData, aes(x=absences)) + 
                   labs(x="The number of absences for Math Course") + 
                   geom_histogram(bins=25, colour="blue", aes(y= ..density.., fill=..count..)) +
                   scale_fill_gradient("count", low="#132B43", high="#56B1F7") +
                   stat_function(fun=dnorm, color="red", args=list(mean=mean(stdMathData$absences, na.rm=TRUE), sd=sd(stdMathData$absences, na.rm=TRUE)))
histogr_ab
boxplot(stdMathData$absences,
        main = "Math course: Absences",
        #xlab = "Number of absences",
        col = "orange",
        border = "brown",
        horizontal = TRUE)

#g1
histogr_g1 <- ggplot(stdMathData, aes(x=g1)) + 
                   labs(x="The first period grade for math course") + 
                   geom_histogram(bins=30, colour="blue", aes(y= ..density.., fill=..count..)) +
                   scale_fill_gradient("count", low="#132B43", high="#56B1F7") +
                   stat_function(fun=dnorm, color="red", args=list(mean=mean(stdMathData$g1, na.rm=TRUE), sd=sd(stdMathData$g1, na.rm=TRUE)))
histogr_g1
boxplot(stdMathData$g1,
        main = "Math course: G1",
        col = "orange",
        border = "brown",
        horizontal = TRUE)

#g2
histogr_g2 <- ggplot(stdMathData, aes(x=g2)) + 
                   labs(x="The second period grade for math course") + 
                   geom_histogram(bins=30, colour="blue", aes(y= ..density.., fill=..count..)) +
                   scale_fill_gradient("count", low="#132B43", high="#56B1F7") +
                   stat_function(fun=dnorm, color="red", args=list(mean=mean(stdMathData$g1, na.rm=TRUE), sd=sd(stdMathData$g1, na.rm=TRUE)))
histogr_g2
boxplot(stdMathData$g1,
        main = "Math course: G2",
        col = "orange",
        border = "brown",
        horizontal = TRUE)

```
```{r normality: qq-plot, echo=TRUE}
#create qq-plot for g1
qqnorm(stdMathData$g1, xlab = "Math course: G1")
qqline(stdMathData$g1, col=2)

qqnorm(stdMathData$g2, xlab="Math course: G2")
qqline(stdMathData$g2, col=2)

qqnorm(stdMathData$absences, xlab="Math course: Number of Absences")
qqline(stdMathData$absences, col=2)

qqnorm(stdMathData$studytime, xlab="Math course: Weekly Study time")
qqline(stdMathData$studytime, col=2)
```
```{r normality:summary statistics g1, echo=TRUE}
#g1 generae summary statistics
pastecs::stat.desc(stdMathData$g1, basic=F)

#skewness and kurtosis
tpskew <- semTools::skew(stdMathData$g1)
tpkurt <- semTools::kurtosis(stdMathData$g1)

tpskew
tpkurt

tpskew[1]/tpskew[2]
tpkurt[1]/tpkurt[2]

mathg1<- abs(scale(stdMathData$g1))
FSA::perc(as.numeric(mathg1), 1.96, "gt")
FSA::perc(as.numeric(mathg1), 3.29, "gt")
```

```{r normality:summary statistics g2, echo=TRUE}
#g2 generae summary statistics
pastecs::stat.desc(stdMathData$g2, basic=F)

#skewness and kurtosis
tpskew <- semTools::skew(stdMathData$g2)
tpkurt <- semTools::kurtosis(stdMathData$g2)

tpskew
tpkurt

tpskew[1]/tpskew[2]
tpkurt[1]/tpkurt[2]

mathg1<- abs(scale(stdMathData$g2))
FSA::perc(as.numeric(mathg1), 1.96, "gt")
FSA::perc(as.numeric(mathg1), 3.29, "gt")
```

```{r normality summary statistics:studytime, echo=TRUE}
#study time generae summary statistics
pastecs::stat.desc(stdMathData$studytime, basic=F)

#skewness and kurtosis
tpskew <- semTools::skew(stdMathData$studytime)

tpkurt <- semTools::kurtosis(stdMathData$studytime)

tpskew
tpkurt
tpskew[1]/tpskew[2]
tpkurt[1]/tpkurt[2]

mathst<- abs(scale(stdMathData$studytime))
FSA::perc(as.numeric(mathst), 1.96, "gt") #greater >= 1.96
FSA::perc(as.numeric(mathst), 3.29, "gt") #greater >= 3.29
```

```{r normality summary statistics:absences, echo=TRUE}
#absences generae summary statistics
pastecs::stat.desc(stdMathData$absences, basic=F)

#skewness and kurtosis
tpskew <- semTools::skew(stdMathData$absences)
tpkurt <- semTools::kurtosis(stdMathData$absences)

#tpskew
#tpkurt

tpskew[1]/tpskew[2]

tpkurt[1]/tpkurt[2]

mathg1<- abs(scale(stdMathData$absences))

#calculate the percentage that are greater than 1.96
FSA::perc(as.numeric(mathg1), 1.96, "gt")
# calculate the percentage that are greater than 3.29
FSA::perc(as.numeric(mathg1), 3.29, "gt")
```

## Correlation

### 1. The correlation between number of absences (absences) and the first period grade (g1)
```{r correlation scatterplot, echo=TRUE}
#show scatterplot of the number of absences and the first period grade

scat_g1 <- ggplot(stdMathData, aes(stdMathData$absences,stdMathData$g1)) +
#show regression line
       geom_point() +
       geom_smooth(method="lm", colour="red", se = F) +
       labs(x = "The Number of Absences", y = "First Period Grade for Math course")
scat_g1
```

```{r correlation pearson, echo=TRUE}
# Conducting Correlation Tests - Pearson, Spearman, Kendall
# Pearson corrlection test --- Parametric test
# Pearson Correlation  (Parametric)
# Spearman Rank Order Correlation or Kendall’s Tau (Non-Parametric)

#1. Pearson corrlation
stats::cor.test(stdMathData$g1,stdMathData$absences, method='pearson')
```

```{r correlation spearman, echo=TRUE}
#2. Spearman corrleation to check the missing data if the data does not fit a normal distribution, then we can use spearman or kurtosis
stats::cor.test(stdMathData$g1, stdMathData$absences, method='spearman')
```

```{r correlation kendall, echo=TRUE}
#3. kendall which handle ties
stats::cor.test(stdMathData$g1, stdMathData$absences, method='kendall')
```

### 2. The correlation between weekly study time (x) and the first period grade (g1) (y)
```{r, echo=TRUE}
#show scatterplot of the number of studytime and the first period grade
scat_st <- ggplot(stdMathData, aes(stdMathData$studytime,stdMathData$g1)) +
#show regression line
       geom_point() +
       geom_smooth(method="lm", colour="red", se = F) +
       labs(x = "Weekly Study time for Math course", y = "First Period Grade for Math course")
scat_st
``` 

```{r orrelation Tests studytime, echo=TRUE}
# Conducting Correlation Tests - Pearson, Spearman, Kendall
# Pearson corrlection test --- Parametric test
# Pearson Correlation  (Parametric)
# Spearman Rank Order Correlation or Kendall’s Tau (Non-Parametric)

#1. Pearson corrlation
stats::cor.test(stdMathData$g1,stdMathData$studytime, method='pearson')
```

```{r, echo=TRUE}
#2. Spearman corrleation to check the missing data if the data does not fit a normal distribution, then we can use spearman or kurtosis
stats::cor.test(stdMathData$g1, stdMathData$studytime, method='spearman')
```

```{r, echo=TRUE}
#3. kendall which handle ties
stats::cor.test(stdMathData$g1, stdMathData$studytime,method='kendall')
```
### 3. The correlation between second period grade (g2)  (y) and the first period grade (g1) (x)
```{r, echo=TRUE}
#show scatterplot of second period grade (g2)  (y) and the first period grade (g1) (x)

scat_g2 <- ggplot(stdMathData, aes(stdMathData$g1, stdMathData$g2)) +
#show regression line
       geom_point() +
       geom_smooth(method="lm", colour="red", se = F) +
       labs(x = "First Period Grade for Math course", y = "Second Period Grade for Math course")
scat_g2
```

```{r, echo=TRUE}
# Conducting Correlation Tests - Pearson, Spearman, Kendall
# Pearson corrlection test --- Parametric test
# Pearson Correlation  (Parametric)
# Spearman Rank Order Correlation or Kendall’s Tau (Non-Parametric)

#1. Pearson corrlation
stats::cor.test(stdMathData$g2, stdMathData$g1, method='pearson')
```

```{r, echo=TRUE}
#2. Spearman corrleation to check the missing data if the data does not fit a normal distribution, then we can use spearman or kurtosis
stats::cor.test(stdMathData$g2, stdMathData$g1, method='spearman')
```

```{r, echo=TRUE}
#3. kendall which handle ties
stats::cor.test(stdMathData$g2, stdMathData$g1, method='kendall')
```
