---
title: 'PSI_Assignment 1: Digital Portfolio: Analyse Part'
author: "Na Li (D19125334)"
---


``` {r setup}

knitr::opts_chunk$set(echo = TRUE)
needed_packages <- c("VIM", "tidyverse","pastecs", "ggplot2", "semTools", "psych", "FSA", "car", "effectsize", "coin", "rstatix", "sjstats", "userfriendlyscience", "stats", "foreign", "gmodels", "lm.beta","stargazer", "lmtest", "DescTools", "nnet", "reshape2", "generalhoslem", "Epi", "arm", "regclass", "olsrr")   
# Extract not installed packages
not_installed <- needed_packages[!(needed_packages %in% installed.packages()[, "Package"])]    
# Install not installed packages
if(length(not_installed)) install.packages(not_installed) 

library(pastecs) #For creating descriptive statistic summaries
library(ggplot2) #For creating histograms with more detail than plot
library(semTools) #For skewness and kurtosis
library(psych)  #For descriptive functions
library(FSA) #For percentage
library(car) # For Levene's test for homogeneity of variance and  test for colinearity of predictors
library(effectsize) #To calculate effect size for t-test
library(VIM)
library(tidyverse)
library(coin) # For Wilcox test (non-parametric)
library(rstatix) # For calculating effect size
library(sjstats) #calculate effect size for t-test
library(userfriendlyscience)
library(stats)
library(foreign) # open SPSS file, I may not use that.
library(gmodels) #For creating histograms with more detail than plot
library(stargazer)#For formatting outputs/tables for regression
library(lm.beta) # to isolate the beta co-efficients for regression

#Multinomial regression
library(lmtest)
library(DescTools)
library(nnet) #Multinomial regression
library(reshape2)
library(generalhoslem) #For test of fit for logistic regression, test assumption of linearity
library(Epi) #ROC Curve
library(arm) #for invlogit calculating predicted probabilities
library(regclass) #For confusion matrix
library(olsrr)
```

```{r initial dataset, echo=TRUE}
getwd()

stdMathData=read.table("student-mat.csv",sep=";",header=TRUE) #read student 
stdPorlagData=read.table("student-por.csv",sep=";",header=TRUE)


stdMergeData=merge(d1,d2,by=c("school","sex","age","address","famsize","Pstatus","Medu","Fedu","Mjob","Fjob","reason","nursery","internet"))

colnames(stdMathData) <- tolower(colnames(stdMathData))
colnames(stdPorlagData) <- tolower(colnames(stdPorlagData))
colnames(stdMergeData) <- tolower(colnames(stdMergeData))
```

# Analyse
## Difference for parametic/non-parametic
### Parametic test (independent t-test/levene's test)
1. absences -- by internet 
```{r 1. absences by internet }
# the variable: the number of absences has generated histogram and summary statistics in th "Explore Data"
# describle the number of absences by internet group
# internet: Internet access at home (binary: yes or no)

# number of absences has been generated histogram and summary statistics in th "Explore Data"
# get descriptive statistics by internet group
psych::describeBy(stdMathData$absences, stdMathData$internet, mat=TRUE)

```
``` {r levenTest and t-test}
# Levene's test for homogenity of variable 
car::leveneTest(absences ~ internet, data=stdMathData)

# Conduct the t-test
# var.equal = True is to specify equal variances and a pooled variance estimate.
res <- stats::t.test(absences ~ internet, var.equal=TRUE, data=stdMathData)
res
```
```{r effect size check: Cohend and Eta}
# statistically significant difference was found
#Cohen's d
# Reporting guideline: Cohen’s d
#0.2=small effect, 0.5=moderate, 0.8 = large

effsize = round((2 * res$statistic)/sqrt(res$parameter), 2)

#using function from effectize package
effectsize::t_to_d(t=res$statistic, res$parameter)
```

```{r effect size check: Eta}
# statistically significant difference was found
# Eta
# reporting guideline: on effect size: 0.01 = small, 0.06 = moderate, 0.14 =large
effes=round((res$statistic*res$statistic)/((res$statistic*res$statistic)+(res$parameter)),3)

effes
```

Report this t-test result:
```
1. Reporting the results with Cohen’s d effect
An independent-samples t-test was conducted to compare the number of absences of Math course for students who have internet at home and those who do not have internet at home. There is a slight Significant difference in the scores for absences was found (M=6.07, SD=8.52 for students who have internet at home, M=3.89, SD=4.25 for students who do not have internet at home), t(393)= -2.027, p-value = 0.043). Cohen's d also indicated a small effect size (-0.20). 

2. Reporting the results with eta squared effect
An independent-samples t-test was conducted to compare the number of absences of Math course for students who have internet at home and those who do not have internet at home. There is a slight Significant difference in the scores for absences was found (M=6.07, SD=8.52 for students who have internet at home, M=3.89, SD=4.25 for students who do not have internet at home), t(393)= -2.027, p-value = 0.043). A small effect size was also indicated by the eta squared value (0.01).
```

2. g3 --- by paid (extra paid classes within the course subject(math) (yes/no)
```{r 2. g3 by paid}
# g3 should be generaged histogram and summary statistics 
ggplot(stdMathData, aes(x=g3))+ 
       labs(x="Final grade for Math course (g3)") + 
       geom_histogram(bins=30, colour="black", aes(y=..density.., fill=..count..)) +             scale_fill_gradient("Count", low="#DCDCDC", high="#7C7C7C") + 
       stat_function(fun=dnorm, color="red",args=list(mean=mean(stdMathData$g3, na.rm=TRUE), sd=sd(stdMathData$g3, na.rm=TRUE)))
# Q-Q plot
qqnorm(stdMathData$g3)
qqline(stdMathData$g3, col=2)
```
```{r g3 normality}
#g3 generae summary statistics
pastecs::stat.desc(stdMathData$g3, basic=F)

#skewness and kurtosis
tpskew <- semTools::skew(stdMathData$g3)
tpkurt <- semTools::kurtosis(stdMathData$g3)

tpskew
tpkurt

tpskew[1]/tpskew[2]
tpkurt[1]/tpkurt[2]

mathg3<- abs(scale(stdMathData$g3))
FSA::perc(as.numeric(mathg3), 1.96, "gt")
FSA::perc(as.numeric(mathg3), 3.29, "gt")
```

```{r 2. g3 by paid}
# get descriptive statistics by paid group
psych::describeBy(stdMathData$g3, stdMathData$paid, mat=TRUE)
```

``` {r levenTest and t-test}
# Levene's test for homogenity of variable 
car::leveneTest(g3 ~ paid, data=stdMathData)

# Conduct the t-test
# var.equal = True is to specify equal variabnces and a pooled variance extimate.
res_g3 <- stats::t.test(g3 ~ paid, var.equal=TRUE, data=stdMathData)
res_g3
```

```{r effect size check: Cohen's d and Eta}
# statistically significant difference was found
#Cohen's d
# Reporting guideline: Cohen’s d
#0.2=small effect, 0.5=moderate, 0.8 = large

effsize = round((2 * res_g3$statistic)/sqrt(res_g3$parameter), 2)

#using function from effectize package
effectsize::t_to_d(t=res_g3$statistic, res_g3$parameter)
```

```{r effect size check: Eta}
# statistically significant difference was found
# Eta
# reporting guideline: on effect size: 0.01 = small, 0.06 = moderate, 0.14 =large
effes=round((res_g3$statistic*res_g3$statistic)/((res_g3$statistic*res_g3$statistic)+(res_g3$parameter)),3)

effes
```
Report this t-test result:
```
1. Reporting the results with Cohen’s d effect
An independent-samples t-test was conducted to compare the final grade (g3) for students who have extra paid classes within the course math subject and those who do not have extra paid classes within the course math subject. There is a slight Significant difference in the scores for the third period grade was found (M=10.92, SD=3.79 for students who have extra paid classes, M=10.00, SD=5.13 for students who do not have extra paid classes), t(393)= -2.03, p-value < 0.05). Cohen's d also indicated a small effect size (-0.21). 

2. Reporting the results with eta squared effect
An independent-samples t-test was conducted to compare the final grade (g3) for students who have extra paid classes within the course math subject and those who do not have extra paid classes within the course math subject. There is a slight Significant difference in the scores for the third period grade was found (M=10.92, SD=3.79 for students who have extra paid classes, M=10.00, SD=5.13 for students who do not have extra paid classes), t(393)= -2.03, p-value < 0.05). A small effect size was also indicated by the eta squared value (0.01).
```

### Non-Parametric test (Mann-Whitney test)
1. Weekly study time (numeric: 1 - <2 hours, 2 - 2 to 5 hours, 3 - 5 to 10 hours, or 4 - >10 hours)
```{r discriptive statistics}
# get descriptive statistics by sex group
by(stdMathData$studytime, stdMathData$sex, median)
by(stdMathData$studytime, stdMathData$sex, IQR)
```

```{r plot histogram}

#split dataset by sex
stdMath_F <- subset(stdMathData, sex="F")
stdMath_M <- subset(stdMathData, set="M")

ggplot(stdMath_F, aes(x=studytime)) +
      ggtitle ("Weekly Study time by sex")+ 
      labs(x="Female: studytime") + 
      geom_histogram(bins=4, colour="black") +
      scale_fill_gradient("Count", low="#DCDCDC", high="#7C7C7C") +
      stat_function(fun=dnorm, color="red",args=list(mean=mean(stdMath_F$studytime,                                                       na.rm=TRUE),sd=sd(stdMath_F$studytime, na.rm=TRUE)))

ggplot(stdMath_M, aes(x=studytime)) +
      ggtitle ("Weekly Study time by sex")+ 
      labs(x="Male: studytime") + 
      geom_histogram(bins=4, colour="black") +
      scale_fill_gradient("Count", low="#DCDCDC", high="#7C7C7C") +
      stat_function(fun=dnorm, color="red",args=list(mean=mean(stdMath_M$studytime,                                                       na.rm=TRUE),sd=sd(stdMath_M$studytime, na.rm=TRUE)))
```
```{r Mann Whitney U (Wilcoxon Test)}
#test for differences on Gender to get U statistic 
stats::wilcox.test(studytime ~ sex, data = stdMathData)
# calculate z-score
coin::wilcox_test(studytime~as.factor(sex), data=stdMathData)
```
```{r effect size}
wilcoeffsize = rstatix::wilcox_effsize(studytime ~ sex, data=stdMathData)
wilcoeffsize
```

```
Reporting the results for the Mann-Whitney test
The weekly study time by female stdMath_F (Mdn=2.00, IQR=1.00) was significantly more study time than male stdMath_M (Mdn=2.00, IQR=1.00), (U=26434.00, z=6.69, p<.001, r=0.34) for Math course sample. 
```

2. Current health status ((numeric: from 1 - very bad to 5 - very good)) by extra-curricular activities (yes/no)
```{r discriptive statistics}
# get descriptive statistics by extra-curricular activities
by(stdMathData$health, stdMathData$activities, median)
by(stdMathData$health, stdMathData$activities, IQR)
```

```{r plot histogram}

#split dataset by activities
stdMathHealAct_N <- subset(stdMathData, activities="no")
stdMathHealAct_Y <- subset(stdMathData, activities="yes")

ggplot(stdMathHealAct_Y, aes(x=health)) +
      ggtitle ("Current health status by activities \n (extra-curricular activities)")+ 
      labs(x="Extra-curricular activities: health") + 
      geom_histogram(bins=5, colour="black") +
      scale_fill_gradient("Count", low="#DCDCDC", high="#7C7C7C") +
      stat_function(fun=dnorm, color="red",args=list(mean=mean(stdMathHealAct_Y$health,                                                       na.rm=TRUE),sd=sd(stdMathHealAct_Y$health, na.rm=TRUE)))

ggplot(stdMathHealAct_N, aes(x=health)) +
      ggtitle("Current health status by nursery \n (attended nursery school)") + 
      labs(x="No extra-curricular activities: health") + 
      geom_histogram(bins=5, colour="black") +
      scale_fill_gradient("Count", low="#DCDCDC", high="#7C7C7C") +
      stat_function(fun=dnorm, color="red",args=list(mean=mean(stdMathHealAct_N$health,                                                       na.rm=TRUE),sd=sd(stdMathHealAct_N$health, na.rm=TRUE)))
```
```{r Mann Whitney U (Wilcoxon Test)}
#test for differences on Gender to get U statistic 
stats::wilcox.test(health ~ activities, data = stdMathData)
# calculate z-score
coin::wilcox_test(health~as.factor(activities), data=stdMathData)
```

```{r effect size}
wilcoeffsize = rstatix::wilcox_effsize(health ~ activities, data=stdMathData)
wilcoeffsize
```

```
Reporting the results for the Mann-Whitney test
The current health status for students who do not have extra-curricular activities stdMathHealAct_N (Mdn=4.00, IQR=2.00) did not differ significantly from those who have extra-curricular activities stdMathHealAct_Y (Mdn=4.00, IQR=2.00), (U=18882.00, z=-0.56, p=0.57, r=0.028). 
```


## Repeat measure for difference
### Parametic test (paired t-test)
G1 and G2 and g3 repeat measure 
```{r g1 and g2 repeat measure}

pastecs::stat.desc(stdMathData$g1, basic=F)
pastecs::stat.desc(stdMathData$g2, basic=F)

# paired t-test g1-g2
res_g12 <- t.test(stdMathData$g1, stdMathData$g2, paired=TRUE)
res_g12
```

```{r effect size check: Eta}
# statistically significant difference was found
# Eta
# reporting guideline: on effect size: .01 = small, .06 = moderate, .14 =large
effes=round((res_g12$statistic*res_g12$statistic)/((res_g12$statistic*res_g12$statistic)+(res_g12$parameter)),3)

effes
```

```
Reporting paired t-test between g1 and g2
A paried-samples t-test was conducted to evaluate the impact of the Math course grades between first period grade and second period grade. There was no statistically significant increase in grade scores from the first period grade (M=10.91, SD=3.32) to the second period grade (M=10.71, SD=3.76), (t(394)=1.96, p=0.05). The mean of increase in grade scores was 0.20 with a 95% confidence interval ranging from -0.12e-3 to 0.39. The eta squared statistic (.01) indicated a small effect size.
```

```{r g2 and g3 repeat measure}
pastecs::stat.desc(stdMathData$g2, basic=F)
pastecs::stat.desc(stdMathData$g3, basic=F)

# paired t-test g1-g2
res_g23 <- t.test(stdMathData$g2, stdMathData$g3, paired=TRUE)
res_g23
```

```{r effect size check: Eta}
# statistically significant difference was found
# Eta
# reporting guideline: on effect size: .01 = small, .06 = moderate, .14 =large
effes=round((res_g23$statistic*res_g23$statistic)/((res_g23$statistic*res_g23$statistic)+(res_g23$parameter)),3)

effes
```

```
Reporting paried t-test between g2 and g3
A paired-samples t-test was conducted to evaluate the impact of the Math course grades between second period grade and third period grade. There was statistically significant increase in grade scores from the second period grade (M=10.71, SD=3.76) to the third period grade (M=10.42, SD=4.58), (t(394)=2.99, p=0.003). The mean of increase in grade scores was 0.30 with a 95% confidence interval ranging from 0.10 to 0.50. The eta squared statistic (.022) indicated a small effect size.
```

### Non-Parametic test --- Math course
```{r dalc and walc by sex repeat measure}
# get subset data by sex
variableint <- c("sex", "dalc", "walc")
stdMath_F <- subset(stdMathData, sex=="F")[variableint]
stdMath_M <- subset(stdMathData, sex=="M")[variableint]

# wilcoxon test should both sub samples are the same length
# randomly choose the data in the female sub-samples
#set.seed(2000)
#stdMath_F <- sample_n(stdMath_F, nrow(stdMath_M))

summary(stdMath_F)

summary(stdMath_M)

```

```{r dalc, walc}
# checking if the statistics dalc and walc are non parametic
# Workday alcohol consumption
ggplot(stdMath_F, aes(x=dalc)) +
      ggtitle ("Workday alcohol consumption")+ 
      labs(x="Workday alcohol consumption: Famle") + 
      geom_histogram(bins=5, colour="black") +
      scale_fill_gradient("Count", low="#DCDCDC", high="#7C7C7C") +
      stat_function(fun=dnorm, color="red",args=list(mean=mean(stdMath_F$dalc,                                                       na.rm=TRUE),sd=sd(stdMath_F$dalc, na.rm=TRUE)))

ggplot(stdMath_F, aes(x=walc)) +
      ggtitle ("Weekend alcohol consumption")+ 
      labs(x="Weekend alcohol consumption: Famle") + 
      geom_histogram(bins=5, colour="black") +
      scale_fill_gradient("Count", low="#DCDCDC", high="#7C7C7C") +
      stat_function(fun=dnorm, color="red",args=list(mean=mean(stdMath_F$walc,                                                       na.rm=TRUE),sd=sd(stdMath_F$walc, na.rm=TRUE)))

# Weekend alcohol consumption
ggplot(stdMath_M, aes(x=dalc)) +
      ggtitle ("Workday alcohol consumption")+ 
      labs(x="Workday alcohol consumption: Famle") + 
      geom_histogram(bins=5, colour="blue") +
      scale_fill_gradient("Count", low="#DCDCDC", high="#7C7C7C") +
      stat_function(fun=dnorm, color="red",args=list(mean=mean(stdMath_M$dalc,                                                       na.rm=TRUE),sd=sd(stdMath_M$dalc, na.rm=TRUE)))

ggplot(stdMath_M, aes(x=walc)) +
      ggtitle ("Weekend alcohol consumption")+ 
      labs(x="Weekend alcohol consumption: Male") + 
      geom_histogram(bins=5, colour="blue") +
      scale_fill_gradient("Count", low="#DCDCDC", high="#7C7C7C") +
      stat_function(fun=dnorm, color="red",args=list(mean=mean(stdMath_M$walc,                                                       na.rm=TRUE),sd=sd(stdMath_M$walc, na.rm=TRUE)))
```
```{r wilcoxon test syntax}
# fmale group: Workday alcohol consumption and weekend alcohol consumption 
stats::wilcox.test(stdMath_F$dalc, stdMath_F$walc, paired=TRUE)
```

```{r wilcoxon test syntax}
# male group: Workday alcohol consumption and weekend alcohol consumption 
stats::wilcox.test(stdMath_M$dalc, stdMath_M$walc, paired=TRUE)
```

```{r wilcoxon test syntax}
coin::wilcox_test(dalc~as.factor(sex), data=stdMathData)
coin::wilcox_test(walc~as.factor(sex), data=stdMathData)

wilcoeffsize1 = rstatix::wilcox_effsize(dalc ~ sex, data=stdMathData)
wilcoeffsize1

wilcoeffsize2 = rstatix::wilcox_effsize(walc ~ sex, data=stdMathData)
wilcoeffsize2

```

```
Reporting the results for the Mann-Whitney test
For female students stdMath_F workday alchol consumption (Mdn=1.00, IQR=0.00) were significantly lower than weekend alchol consumption (Mdn=2.00, IQR=2.00), V = 79.50, p < .001. For male students stdMath_M, workday alchol consumption (Mdn=1.00, IQR=1.00) were lower than weekend alchol consumption (Mdn=3, IQR=3), V=72, p < .001.
```

### Non-Parametic test ---language course
```{r dalc and walc by sex repeat measure}
# get subset data by sex
variableint <- c("sex", "dalc", "walc")
stdPorlag_F <- subset(stdPorlagData, sex=="F")[variableint]
stdPorlag_M <- subset(stdPorlagData, sex=="M")[variableint]

# wilcoxon test should both sub samples are the same length
# randomly choose the data in the female sub-samples
#set.seed(2000)
#stdMath_F <- sample_n(stdMath_F, nrow(stdMath_M))

summary(stdPorlag_F)
pastecs::stat.desc(stdPorlag_F, basic=F)
summary(stdPorlag_M)
pastecs::stat.desc(stdPorlag_M, basic=F)

```

```{r dalc, walc}
# checking if the statistics dalc and walc are non parametic
# Workday alcohol consumption
ggplot(stdPorlag_F, aes(x=dalc)) +
      ggtitle ("Workday alcohol consumption")+ 
      labs(x="Workday alcohol consumption: Famle") + 
      geom_histogram(bins=5, colour="black") +
      scale_fill_gradient("Count", low="#DCDCDC", high="#7C7C7C") +
      stat_function(fun=dnorm, color="red",args=list(mean=mean(stdMath_F$dalc,                                                       na.rm=TRUE),sd=sd(stdMath_F$dalc, na.rm=TRUE)))

ggplot(stdPorlag_F, aes(x=walc)) +
      ggtitle ("Weekend alcohol consumption")+ 
      labs(x="Weekend alcohol consumption: Famle") + 
      geom_histogram(bins=5, colour="black") +
      scale_fill_gradient("Count", low="#DCDCDC", high="#7C7C7C") +
      stat_function(fun=dnorm, color="red",args=list(mean=mean(stdMath_F$walc,                                                       na.rm=TRUE),sd=sd(stdMath_F$walc, na.rm=TRUE)))

# Weekend alcohol consumption
ggplot(stdPorlag_M, aes(x=dalc)) +
      ggtitle ("Workday alcohol consumption")+ 
      labs(x="Workday alcohol consumption: Male") + 
      geom_histogram(bins=5, colour="blue") +
      scale_fill_gradient("Count", low="#DCDCDC", high="#7C7C7C") +
      stat_function(fun=dnorm, color="red",args=list(mean=mean(stdMath_M$dalc,                                                       na.rm=TRUE),sd=sd(stdMath_M$dalc, na.rm=TRUE)))

ggplot(stdPorlag_M, aes(x=walc)) +
      ggtitle ("Weekend alcohol consumption")+ 
      labs(x="Weekend alcohol consumption: Male") + 
      geom_histogram(bins=5, colour="blue") +
      scale_fill_gradient("Count", low="#DCDCDC", high="#7C7C7C") +
      stat_function(fun=dnorm, color="red",args=list(mean=mean(stdMath_M$walc,                                                       na.rm=TRUE),sd=sd(stdMath_M$walc, na.rm=TRUE)))
```
```{r wilcoxon test syntax}
# fmale group: Workday alcohol consumption and weekend alcohol consumption 
stats::wilcox.test(stdPorlag_F$dalc, stdPorlag_F$walc, paired=TRUE)
```

```{r wilcoxon test syntax}
# male group: Workday alcohol consumption and weekend alcohol consumption 
stats::wilcox.test(stdPorlag_M$dalc, stdPorlag_M$walc, paired=TRUE)
```

```{r}
stats::wilcox.test(stdPorlagData$dalc, stdPorlagData$walc, paired=TRUE)
```

```{r wilcoxon test syntax}
coin::wilcox_test(dalc~as.factor(sex), data=stdPorlagData)
coin::wilcox_test(walc~as.factor(sex), data=stdPorlagData)

wilcoeffsize1 = rstatix::wilcox_effsize(dalc ~ sex, data=stdPorlagData)
wilcoeffsize1

wilcoeffsize2 = rstatix::wilcox_effsize(walc ~ sex, data=stdPorlagData)
wilcoeffsize2

```

```
Reporting the results for the Mann-Whitney test
    For female students stdPorlagData_F workday alcohol consumption (Mdn=1.00, IQR=0.00) were significantly lower than weekend alcohol consumption (Mdn=2.00, IQR=2.00), V = 884.5, p < 0.001, r=0.28. 
   For male students stdPorlagData_M, workday alcohol consumption (Mdn=1.00, IQR=1.00) were lower than weekend alcohol consumption (Mdn=3, IQR=2), V=292, p < 0.001, r=0.30. 
     What's more, For workday alcohol consumption for both female and male students was significantly lower than weekend alcohol consumption for both female and male students(V = 2241.5, p-value < 0.001). 
```
#### Differences Nominal 
for two independent variables in Math course.
Sex - schoolup
Questioin: “Is there a difference between boys and girls in terms of whether they have extra educational support?”
H0:  There is no difference
HA: There is a difference

```{r Contingency table, echo=TRUE}
#CrossTable(predictor, outcome, fisher = TRUE, chisq = TRUE, expected = TRUE)
gmodels::CrossTable(stdMathData$sex, stdMathData$schoolsup, fisher = TRUE, chisq = TRUE, expected = TRUE, sresid = TRUE, format = "SPSS")
```
```{r}
# simplistic way of doing Chi-Square

#Create your contingency table
contintable<-xtabs(~schoolsup+sex, data=stdMathData)

ctest<-stats::chisq.test(contintable, correct=TRUE)#chi square test
#correct=TRUE to get Yates correction needed for 2x2 table

ctest#will give you the details of the test statistic and p-value

cat("\n ****************Expected frequencies:****************\n") 
ctest$expected#expected frequencies
cat("\n ****************Observed frequencies:****************\n") 
ctest$observed#observed frequencies
cat("\n *****p-value:Pearson's Chi-squared test with Yates' continuity correction *****\n") 
ctest$p.value
```
```{r effect size}
#Calculate effect size
sjstats::phi(contintable)
sjstats::cramer(contintable)
```
```
report Chi-Square Test for Independence

A Chi-Square test for independence (with Yates’ Continuity Correction) indicated significant association between gender and extra educational support, χ2(1,n=395)=6.75,p<0.05, phi=0.14).

```

### Difference for more than two groups 
#### Parametic
1. studytime weekly study time (numeric: 1 - <2 hours, 2 - 2 to 5 hours, 3 - 5 to 10 hours, or 4 - >10 hours) group for final grade (g1)
```{r normality, echo=TRUE}
# g3 Examine/Asses Normality

gg_g1 <- ggplot(stdPorlagData, aes(x=g1)) +
         ggtitle("Figure 1 - Histogram for first period grade for Portuguese language course") + 
         labs(x="First period grade for Portuguese Language course") + 

#manage binwidth and colours
          geom_histogram(binwidth=2, colour="black", aes(y=..density.., fill=..count..)) +
          scale_fill_gradient("Count") + 

#adding a normal curve
#use stat_function to compute a normalized score for each value of g1
#pass the mean and standard deviation
#use the na.rm parameter to say how missing values are handled
         stat_function(fun=dnorm, color="red",args=list(mean=mean(stdPorlagData$g1, na.rm=TRUE), sd=sd(stdPorlagData$g1, na.rm=TRUE)))

#to display the graph request the contents of the variable be shown
gg_g1
```
```{r qq plot for g3, echo=TRUE}
qqnorm(stdPorlagData$g1, main = "Figure 2 - QQ Plot for first period grade of Portuguese Language course")
qqline(stdPorlagData$g1, col = 2) # show a trend line on the plot.
```
```{r statistic description, echo=TRUE}
#summary(stdPorlagData)
pastecs::stat.desc(stdPorlagData$g1, basic=F)
```
```{r calculate skew and kurtosis for g3, echo=TRUE}
tpskew <- semTools::skew(stdPorlagData$g1)
tpkurt <- semTools::kurtosis(stdPorlagData$g1)
#skew calculate
tpskew[1]/tpskew[2]
#kurtosis calculate
tpkurt[1]/tpkurt[2]
```
```{r standarised scores, echo=TRUE}
# Calculate the percentage of standardized scores that are greated than 1.96
ztotsas <- abs(scale(stdPorlagData$g1))
FSA::perc(as.numeric(ztotsas), 1.96, "gt") # 'gt' = greater than 1.96, "leq" = equal, "lt"=less than
FSA::perc(as.numeric(ztotsas), 3.29, "gt")
```
```
Report assessment of normality for g1 (first period grade of Portuguese Language)

The first period grades of Portuguess Language course was assess for normality. Visual inspection of the histogram and QQ-Plot (refer Fig. 1 and Fig 2) identified some issues with skewness and kurtosis. The standardised score for skewness (-0.02884671) can be considered acceptable and the standarised score for kurtosis (0.1905247) was inside the accpetable range. Moreover more 99% of standarised scores for final grades fall within the bounds of +/- 3.29, the data can be considered to approximate a normal distribution (m=22.12, sd=2.75, n=649)

```

2. Differences more than 2 groups (ANOVA)
Describle first peroid grades by group (studytime: group1, group 2, group 3, and group 4)
```{r descriptive statistics by group, echo=TRUE}
psych::describeBy(stdPorlagData$g1, stdPorlagData$studytime, mat=TRUE)
```
```{r homogeneity check, echo=TRUE}
#conduct Bartlett's test for homogeneity of variance in library car
stats::bartlett.test(g1 ~ studytime, data=stdPorlagData)
```
``` {r Conduct one-way ANOVA, echo=TRUE}
#p-value is > 0.05 so the result is not statistically significant so we can assume homogeneity
userfriendlyscience::oneway(as.factor(stdPorlagData$studytime), y=stdPorlagData$g1, posthoc="Tukey")
```
```{r effect size and reporting}
# for reporting
res2<-stats::aov(g3 ~ studytime, data = stdPorlagData)
res2

summary(res2) #get F statistic and p-value

#Get the F statistic into a variable
fstat<-summary(res2)[[1]][["F value"]][[1]]
fstat
#Get the p value
aovpvalue<-summary(res2)[[1]][["Pr(>F)"]][[1]]
aovpvalue

#Calculate effect
aoveta<-sjstats::eta_sq(res2)[2]
aoveta
#In the report we are using the res2 variable to retrieve the degrees of freedom
#and the eta_sq function from the sjstats package to calculate the effect
```
```
Reporting the results with eta squared effect
An one-way between-groups analysis of variance (ANOVA) was conducted to explore the impact of weekly study time on the first period grades for Portuguese Language course. Participants were divided into four groups according to their weekly study time (Group 1: < 2 hours; Group 2: 2 to 5 hours; Group 3: 5 to 10 hours; Group 4: > 10 hours). There was a statistically significant difference at the p < 0.001 level in first period grades for the four weekly study time groups: (F(3, 645) = 16.24), p<0.001.Despite reaching statistical significance, the actual difference in mean scores between groups was small. A small effect size was also indicated by the eta squared value (0.06). Post-hoc comparisons using the Tukey HSD test indicated that the mean scores for Group 1 (M=10.50, SD=2.57) was significantly difference to that for Group 2 (M=11.54, SD=2.74), Group 3 (M=12.42, SD=2.46), Group 4(M=12.77, SD=2.94); the mean scores for Group 2 was significantly difference to that for other three groups; Group 3 did not differ significantly for Group 4.

```

#### No Parametic
1. Kruskal-Wallis (Non-Parametric)
2. Friedmans ANOVA (non-parametric) for repeated measures

##### Kruska-wallis test
Nominal --- Guardian (student's guardian (mother, father or other))
Ordinal --- health (current health status (from 1 - very bad to 5 - very good))
group by guardian (3 groups)
```{r statistic descriptive}
psych::describeBy(as.numeric(stdPorlagData$health), factor(stdPorlagData$guardian),mat=TRUE)

```
``` {r kruskal - Wallis rank sum test}
#run a kruskal wallis test
stats::kruskal.test(health~guardian,data=stdPorlagData)
```
```{r post-hot test for between groups}
tmp_bonferroni<-FSA::dunnTest(x=as.numeric(stdPorlagData$health), g=factor(stdPorlagData$guardian), method="bonferroni") 
print(tmp_bonferroni, dunn.test.results = TRUE)
```
```{r effect size}
# bootstrapping
rstatix::kruskal_effsize(stdPorlagData, health~guardian, ci = FALSE, conf.level = 0.95, ci.type = "perc", nboot = 1000)
```
```
An Kruskal-Wallis rank sum was conducted to to explore the impact of guardian on levels of students health. Participants were divided into three groups according to their guardian (Group 1: Father; Group 2: Mother; Group 3: other). No statistically significant difference in the scores for the levels of health (Mdn=4.00, M=3.50, SD=1.42 for Group 1, Mdn= 4.00, M=3.54, SD=1.46 for Group 2, Mdn=4, M=3.59, SD=1.41 for Group 3) (H = 0.233, p=0.89). A very small effect size was also indicated by the kruskal value (-0.027e-102).
```


##### Friedmans ANOVA (parametric) for repeated measures
G1 G2 G3 for each students ranks

```{r }
Responses <- na.omit(with(stdPorlagData, cbind(g1, g2, g3)))
cat("\nMedians:\n") 
#print out the medians of each of our variables
print(apply(Responses, 2, median)) 
```
``` {r friedman test}
stats::friedman.test(Responses)
```
```
Report Firedman test

The three grades of participants for Portuguese Language course has strong significantly change over the three periods grades, χ2(2) = 6.75, p < 0.001.
```



